---
title: "Cribbage hand optimiser"
author: "Dean Findlay"
date: "24 November 2015"
output: html_document
---

## Initialise the data frame 


The array will contain 52 rows with columns of value, suit and split (bool to show if this was cut)  

The cards will be arranged:  
1,1,1,1,2,2,2,2,3... 
c,d,h,s,c,d,h,s...
0,0,0,0,0...




## Choose six random cards from the pack as well as the cut card (to be used later)  

Random cards will be used first then changed to chosen cards once up and running   


```{r}
# package this up to a function

# create the pack 
pack <- data.frame( value = rep(1:13, each = 4),
                    suit = rep(c("c","d","h","s"), 13),
                    cut = FALSE,
                    stringsAsFactors = FALSE) 
# create a vector with the position of the 6 cards in the pack
randoms <- sample(1:52, 6)
# use this vector to subset from the pack
hand <- pack[randoms,]
# remove the chosen cards from the pack
pack <- pack[-randoms,]

createRandomHand<- function(pack,hand)
{
  pack <- data.frame( value = rep(1:13, each = 4),
                      suit = rep(c("c","d","h","s"), 13),
                      cut = FALSE,
                      stringsAsFactors = FALSE) 
  # create a vector with the position of the 6 cards in the pack
  randoms <- sample(1:52, 6)
  # use this vector to subset from the pack -assign in parents scope
  hand <<- pack[randoms,]
  # remove the chosen cards from the pack - assign in parents scope
  pack <<- pack[-randoms,]
}

# change the 5th card to the cut card and remove 6th - only for use while testing
hand[5,3]<- TRUE
hand<-hand[1:5,]

# order the hand by value and suit
hand <- hand[order(hand[,1], hand[,2]),]

```


```{r}
# pick the cut card location
cutLoc <- sample(1:46,1)
# create the card
cutCard <- pack[cutLoc,]
# change the 'cut' field to TRUE
cutCard[,"cut"] <- TRUE
# remove it from the pack
pack <- pack[-cutLoc,]


```



## define the functions that score the hand

the hands are scored using the cut card also although this card can sometimes act differently than cards in the hand as will be discussed.  
All the functions are passed a data frame with the hand and the cut card in it ordered by value and suit.  


!!!!!!!!!!!!!!!!!!!!!!!! move the helper function definitions such as moreCards and cycleRuns into their own bit so they don't get re-defined everytime the main function using them gets called !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

### 15's  

every combination of cards that adds up to 15 scores 2 points. Cards can be used multiple times but the combinations of all sets must be unique i.e. a hand with one 8 and two 7's can have to sets of 15's one for each 7 with the 8.  

I can't think of any elegant ways to achieve this so a brute force method may have to be used, not too bad though as there are not too many combinations of cards to cycle through.  
also as the hand is ordered then if any combination of 2 cards are > 15 the remaining combinations will also be > 15 so the function can end.  
```{r}

count15s<- function(hand)
{

  # define recursive function to keep checking against the rest of the hand
  # vect = vector of the values in the hand with face cards = 10
  # i = current iteration in the loop
  # cntr = current No of cards from from i that is being used
  # total = the current total of the values of cards currently in use
  # points = current amount of points scored by the hand
  moreCards<- function(vect, i, cntr, total, points, startCntr)
  {
    # check to see if trying to look at a 6th card
    if((cntr+i) > 5)
    {
      # check to see if start counter at end
      # yes means leave function as all cards for this i looked at
      if((startCntr+i) > 5)
      {
        return(points)
      }
      # no means add one to startCntr and start carry on with function 
      else
    {
      startCntr<- startCntr + 1
      # make total = i.value
      total<- vect[i]
      # make cntr = startCntr
      cntr = startCntr
      # call morecards 
      moreCards(vect, i, cntr, total, points, startCntr)
    }
  }
  # ok to carry on with function
  else
  {
    # add the next card to the total
    total<- total + vect[i+cntr]
    
    # check to see if total greater than 15 
    if(total > 15)
    {
      # check to see if there is a gap between the startCntr and cntr > 1
      # this will dictate whether previous cards need to be removed and later
      # ones added to ensure all combinations checked
      
      # only interested in finding combination using i & startCntr in them as other will be taken care
      # of later on in the recursion
      
      # if true               !!!!!!!!!!!!!!!!!!!! this is the problem bit !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
      # get this bit to check for whatever and do all the stuff required to get it to the startCntr++ bit
      # then get rid of the else statement at the end and just have it run through that bit every time
      
      if(cntr - startCntr > 1)
      {
        
        # if diff = 3
        if(cntr - startCntr == 3)
        {
          # just check the 3 combinations available for = 15
          if(vect[i]+vect[i+startCntr]+vect[i+cntr] == 15)
          {
            points<- points + 2
          }
          
          if(vect[i]+vect[i+startCntr]+ vect[i+cntr-2]+vect[i+cntr] == 15)
          {
            points<- points + 2
          }
          
          if(vect[i]+vect[i+startCntr]+ vect[i+cntr-1]+vect[i+cntr] == 15)
          {
            points<- points + 2
          }  
        }
        # else diff = 2
        else
        {
          # remove the value of the previous card from the total
          total<- total - vect[i+cntr-1]
          # check to see if the total is 15 and add points if true
          if(total == 15)
          {
            points<- points + 2
          }
          # check to see if there is a card further in the hand that needs adding in
          # i.e a hand of vals 4,4,4,4,7 would be > 15 by card 4
          # thisd can only happen when i=1 startCntr=1 & cntr = 3
          if(i+cntr < 5)
          {
            # just check the 3 combinations available for = 15
            if(vect[i]+vect[i+startCntr]+vect[i+cntr+1] == 15)
            {
              points<- points + 2
            }
            
            if(vect[i]+vect[i+startCntr]+ vect[i+cntr-1]+vect[i+cntr+1] == 15)
            {
              points<- points + 2
            }
            
            if(vect[i]+vect[i+startCntr]+ vect[i+cntr]+vect[i+cntr+1] == 15)
            {
              points<- points + 2
            }
          }
        }
      }
      
      # move startCntr onto the next card to check further combinations
      startCntr<- startCntr + 1
      # make total = i.value
      total<- vect[i]
      # make cntr = startCntr
      cntr = startCntr
      # call morecards 
      moreCards(vect, i, cntr, total, points, startCntr)
      
      
    }
    #else if check to see if total = 15 if true then points need adding and
    # the last card needs removing from the total and the next card adding
    # in case it is of the same value therfore needing more ponts
    else if(total == 15)
    {
      
      # check for diff in startCntr and cntr to see if any other combos are needed to be checked
      # also if startCntr = cntr then further combos will be taken care of later on in the recursion
      
      if(startCntr == cntr)
      {
        points<- points + 2
        # move startCntr onto the next card to check further combinations
        startCntr<- startCntr + 1
        # make total = i.value
        total<- vect[i]
        # make cntr = startCntr
        cntr = startCntr
        # call morecards 
        moreCards(vect, i, cntr, total, points, startCntr)
      }
      else
      {
        points<- points + 2
        total <- total - vect[i+cntr]
        cntr<- cntr + 1
        moreCards(vect, i, cntr, total, points,startCntr)
      }
      
    }
    #else the total is less than 15 and more cards need adding
    else
    {
      cntr<- cntr + 1
      moreCards(vect, i, cntr, total, points, startCntr)
    }
  }
}

  # initialise points
  points <- 0
  
  #obtain values from hand
  vect<- hand[,1]
  # change the values of any 'face' cards from 11, 12 or 13 to 10
  vect[vect>10]<-10

  for(i in 1:4)
  {
    startCntr <- i
    cntr<- 1
    total <- vect[i]
    
#    # if combo of 2 cards greater than 15 exit the loop
#    if(total + vect[i+a] > 15)
#    {
#      return(points)
#    }
  
    points<- moreCards(vect, i, cntr, total, points, startCntr)
  }
  
  return(points)
}

```


### Pairs

every combination of pairs will be awarded 2 points. each card can be used multiple times but the combination of both cards must be unique i.e. a hand with 3 7's will score 6 points; 2 for each combination of pairs.  


```{r}

# iterates through finding duplicates multpile times in order to account for
# 3 and four of a kind adding the extra points where applicable
pair <- function(hand){
  points<- 0
  
  kind2 <- hand[duplicated(hand[,1]),1]
  points = points + length(kind2) * 2
  kind3 <- kind2[duplicated(kind2)]
  points = points + length(kind3) * 2
  kind4 <- kind3[duplicated(kind3)]
  points = points + length(kind4) * 2
  
  return(points)
}

```


### Flush  

To score a flush each card in the hand must be of the same suit and will score 4 points if this condition is met. If the cut card is also the same suit then an extra point willbe awarded.  

```{r}

flushScorer <- function(hand){
  # reduce the hand to unique suits, if there is only 1 suit then add 4 to points
  if (length(unique(hand[hand$cut == FALSE,2])) == 1){
    points <- 4
    # if the suit of the cut card matches the suit of the cards in the hand
    # make points = 5
    if(unique(hand[hand$cut == TRUE,2]) == unique(hand[hand$cut == FALSE,2])){
      points <- 5
    }
    # if there is more than one unique suit in the pack then points = 0
  } else { 
    points <- 0}
  
  return(points)
}


```



### Runs  

Runs are any combination of 3 or more cards in sequential order and score one point for each card. cards can be used multiple times but the combination to make the runs must be unique i.e. a hand that has the cards 2h, 3h, 3s, 4h, 9h can have 2 runs of:  
2h, 3h, 4h,  
2h, 3s, 4h  

A further rule is that runs must be the maximum length that they can be i.e. a run of 1,2,3,4 cannot be split into 3 runs of:  
1,2,3  
2,3,4  
1,2,3,4  

```{r}

countRuns<- function(hand)
{
  # vect is a vector of the values in the hand
  # cntr is the current position being examined - initially set at 2
  # currentVal sent in initially as the first value in the hand
  # runLength states how many cards the run has been going on for
  # multipleLength states how many preceding cards have had the same value - initially set as 0
  # endMultiple is a bool that states whether multiple cards have alreay been present in hand on the run
  cycleRuns<- function(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
  {
    
    # check that cntr is not greater than 5 so its not looking for a 6th card
    if(cntr < 6)
    {
      # is the value of cntr within 1 of currentVal continue function if not exit as no other
      if(vect[cntr] < currentVal+2)
      {
        # is the value the same as current val
        # if yes then set onMultiple to TRUE and multipleLength to 1
        if(vect[cntr] == currentVal)
        {
          # check to see if this is 3rd or 4th card in a row of same val
          if(onMultiple)
          {
            multipleLength<- multipleLength + 1
            currentVal<- vect[cntr]
            cntr<- cntr + 1
            cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
          }
          # otherwise it is a new multiple and the 2 cards need adding to multiple length
          else
          {
            multipleLength<- multipleLength + 2
            onMultiple<- TRUE
            currentVal<- vect[cntr]
            cntr<- cntr + 1
            cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
          }
        }
        # the card is on a potential run
        else
        {
          runLength<- runLength +1
          onMultiple<- FALSE
          currentVal<- vect[cntr]
          cntr<- cntr + 1
          cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
        }
      }
      # else not part of a run
      # check that a run of 3 or greater has not ended as their will be no more runs after this
      #   and the points will need to be calculated
      else if(runLength > 2)
      {
        #if multipleLenght = 0 change it to 1 so it can be multiplied with runLength
        if(multipleLength == 0)
        {
          multipleLength<- 1
        }
        
        return(runLength * multipleLength)
      }
      # reset counters and carry on
      else
      {
        onMultiple<- FALSE
        runLength<- 1
        multipleLength<- 0
        currentVal<- vect[cntr]
        cntr<- cntr + 1
        cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
      }
    }
    # check to see if the hand ended on a run
    else if(runLength > 2)
    {
      #if multipleLenght = 0 change it to 1 so it can be multiplied with runLength
      if(multipleLength == 0)
      {
        multipleLength<- 1
      }
      
      return(runLength * multipleLength)
    }
    else
    {
      return(0)
    }
  }
  
  # get values of the hand
  vect<- hand[,1]
  
  # initiate cycleRuns()
  points<- cycleRuns(vect, 2, vect[1],1, 0,FALSE)
  
  return(points)
}


```


### One for his nob

if a hand has a jack that has the same suit as the cut card then an extra point is awarded

```{r}

nobs<- function(hand){
  points <- 0
  # returns the suit of any cards from hand (not cut) that are jacks
  jacks <- hand[hand$value==11 & hand$cut==FALSE,2]

  # if length of jacks is greater than zero get the suit of the cut card
  # join jacks and cut card together and sum the check for duplicates

  if (length(jacks) > 0){
    # returns the suit of the cut card
    cut <- hand[hand$cut==TRUE,2]
    suits <- c(jacks,cut)
    points <- sum(duplicated(suits))
  }

  return(points)
  
}

```


## Create a function that takes the hand, containing the cut card, and calls all point scoring functions to calculate the score

```{r}

calculateScore<- function(hand)
{
  points<- count15s(hand)
  points<- points + countRuns(hand)
  points<- points + flushScorer(hand)
  points<- points + pair(hand)
  points<- points + nobs(hand)
  return(points)
}

```


## Loop through the different combinations of discarded cards, calculating the points and creating a data frame to hold results

The function will need to run 15 times in order to choose every combination of discarded cards  
a data frame will need to be created before any loops to hold the results for each set of discarded cards and the points they score for each cut card.


function will take a hand of 6 cards, remove 2 at a time covering all combinations and calculate the score for each remaining hand with all remaining cards in the pack as the cut card.  


either one data frame for all or one for each set of discarded cards?


```{r}





```

 
### this works through the discard positions
a <- 1:6

for(i in 1:5)
{
  for(j in i+1:(6-i))
  {
    print(c(i,j))
    # use index i and j to discard cards and create hand
    
    # loop to go through each remaining card in the pack as the cut card and work out points
    
      # function to calculate points here
  }
}







## Displaying the output

To be of any use the output needs to display the best hands, but what makes a hand the best?  

Metrics could be any or a combination of the following:
- Maximum score  
- Mean score  
- Quartile ranges  
- Mode  
- Sum of all scores  


With the 'best' defined the output needs to be displayed  

Options could be:
- Display all 15 as bar graphs  
- Display the top 5 as bar graphs  
- Tabular summary of all or top hands with a selection of bar graphs  

When ported to Shiny it may be possible to display the tabular data with an option to dsiplay specific graphs or a range of the top ones  













