---
title: "FunctionTesting"
author: "Dean Findlay"
date: "3 December 2015"
output: html_document
---

This area is to be used to test the functions.  
expected inputs and outputs should be stated and functions only moved into main program until it works as expected


# moreCards

## expected input and output  

moreCards(vect=c(5,10,10,10,10),i=1,cntr=1,total=5,points=0,startCntr=1) # == 8 y
moreCards(vect=c(1,3,9,10,10),i=1,cntr=1,total=1,points=0,startCntr=1) # == 0 y
moreCards(vect=c(4,4,4,7,9),i=1,cntr=1,total=4,points=0,startCntr=1) # == 4 y
moreCards(vect=c(4,4,4,7,9),i=2,cntr=1,total=4,points=4,startCntr=1) # == 6 y   (this is starting on the second card)
moreCards(vect=c(4,4,4,4,7),i=1,cntr=1,total=4,points=0,startCntr=1) # == 6 y
moreCards(vect=c(4,4,4,7,7),i=1,cntr=1,total=4,points=0,startCntr=1) # == 8 y
moreCards(vect=c(2,2,3,4,7),i=1,cntr=1,total=2,points=0,startCntr=1) # == 2 y
moreCards(vect=c(2,2,4,4,7),i=1,cntr=1,total=2,points=0,startCntr=1) # == 4 y
moreCards(vect=c(5,5,5,5,10),i=1,cntr=1,total=2,points=0,startCntr=1) # == 6 y
moreCards(vect=c(2,3,10,10,10),i=1,cntr=1,total=2,points=0,startCntr=1) # == 6 y
moreCards(vect=c(5,5,5,5,10),i=1,cntr=1,total=2,points=0,startCntr=1) # == 6 y


```{r}

moreCards<- function(vect, i, cntr, total, points, startCntr)
{
  # check to see if trying to look at a 6th card
  if((cntr+i) > 5)
  {
    # check to see if start counter at end
    # yes means leave function as all cards for this i looked at
    if((startCntr+i) > 5)
    {
      return(points)
    }
    # no means add one to startCntr and start carry on with function 
    else
    {
      startCntr<- startCntr + 1
      # make total = i.value
      total<- vect[i]
      # make cntr = startCntr
      cntr = startCntr
      # call morecards 
      moreCards(vect, i, cntr, total, points, startCntr)
    }
  }
  # ok to carry on with function
  else
  {
    # add the next card to the total
    total<- total + vect[i+cntr]
    
    # check to see if total greater than 15 
    if(total > 15)
    {
      # check to see if there is a gap between the startCntr and cntr > 1
      # this will dictate whether previous cards need to be removed and later
      # ones added to ensure all combinations checked
      
      # only interested in finding combination using i & startCntr in them as other will be taken care
      # of later on in the recursion
      
      # if true               !!!!!!!!!!!!!!!!!!!! this is the problem bit !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
      # get this bit to check for whatever and do all the stuff required to get it to the startCntr++ bit
      # then get rid of the else statement at the end and just have it run through that bit every time
      
      if(cntr - startCntr > 1)
      {
        
        # if diff = 3
        if(cntr - startCntr == 3)
        {
          # just check the 3 combinations available for = 15
          if(vect[i]+vect[i+startCntr]+vect[i+cntr] == 15)
          {
            points<- points + 2
          }
          
          if(vect[i]+vect[i+startCntr]+ vect[i+cntr-2]+vect[i+cntr] == 15)
          {
            points<- points + 2
          }
          
          if(vect[i]+vect[i+startCntr]+ vect[i+cntr-1]+vect[i+cntr] == 15)
          {
            points<- points + 2
          }  
        }
        # else diff = 2
        else
        {
          # remove the value of the previous card from the total
          total<- total - vect[i+cntr-1]
          # check to see if the total is 15 and add points if true
          if(total == 15)
          {
            points<- points + 2
          }
          # check to see if there is a card further in the hand that needs adding in
          # i.e a hand of vals 4,4,4,4,7 would be > 15 by card 4
          # thisd can only happen when i=1 startCntr=1 & cntr = 3
          if(i+cntr < 5)
          {
            # just check the 3 combinations available for = 15
            if(vect[i]+vect[i+startCntr]+vect[i+cntr+1] == 15)
            {
              points<- points + 2
            }
            
            if(vect[i]+vect[i+startCntr]+ vect[i+cntr-1]+vect[i+cntr+1] == 15)
            {
              points<- points + 2
            }
            
            if(vect[i]+vect[i+startCntr]+ vect[i+cntr]+vect[i+cntr+1] == 15)
            {
              points<- points + 2
            }
          }
        }
      }
      
      # move startCntr onto the next card to check further combinations
      startCntr<- startCntr + 1
      # make total = i.value
      total<- vect[i]
      # make cntr = startCntr
      cntr = startCntr
      # call morecards 
      moreCards(vect, i, cntr, total, points, startCntr)
      
      
    }
    #else if check to see if total = 15 if true then points need adding and
    # the last card needs removing from the total and the next card adding
    # in case it is of the same value therfore needing more ponts
    else if(total == 15)
    {
      
      # check for diff in startCntr and cntr to see if any other combos are needed to be checked
      # also if startCntr = cntr then further combos will be taken care of later on in the recursion
      
      if(startCntr == cntr)
      {
        points<- points + 2
        # move startCntr onto the next card to check further combinations
        startCntr<- startCntr + 1
        # make total = i.value
        total<- vect[i]
        # make cntr = startCntr
        cntr = startCntr
        # call morecards 
        moreCards(vect, i, cntr, total, points, startCntr)
      }
      else
      {
        points<- points + 2
        total <- total - vect[i+cntr]
        cntr<- cntr + 1
        moreCards(vect, i, cntr, total, points,startCntr)
      }
      
    }
    #else the total is less than 15 and more cards need adding
    else
    {
      cntr<- cntr + 1
      moreCards(vect, i, cntr, total, points, startCntr)
    }
  }
}



```


# cycleRuns

## expected input and output

cycleRuns(vect=c(1,3,5,7,9),cntr=2, currentVal=1, runLength=1, multipleLength=0, onMultiple=FALSE) # == 0 yes
cycleRuns(vect=c(1,2,3,4,5),cntr=2, currentVal=1, runLength=1, multipleLength=0, onMultiple=FALSE) # == 5 yes
cycleRuns(vect=c(1,1,2,3,4),cntr=2, currentVal=1, runLength=1, multipleLength=0, onMultiple=FALSE) # == 8 yes
cycleRuns(vect=c(1,1,2,2,3),cntr=2, currentVal=1, runLength=1, multipleLength=0, onMultiple=FALSE) # == 12 yes
cycleRuns(vect=c(1,2,2,2,3),cntr=2, currentVal=1, runLength=1, multipleLength=0, onMultiple=FALSE) # == 9 yes
cycleRuns(vect=c(1,1,2,3,4),cntr=2, currentVal=1, runLength=1, multipleLength=0, onMultiple=FALSE) # == 8 yes


```{r}

# currentVal sent in initially as the first value in the hand

# endMultiple is a bool that states whether multiple cards have alreay been present in hand on the run
# runLength states how many cards the run has been going on for
# multipleLength states how many preceding cards have had the same value
# runStart is the index that a run has started

# the value of the first card is passed in as the current value and cntr will initially be set at 2
cycleRuns<- function(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
{
  
  # check that cntr is not greater than 5 so its not looking for a 6th card
  if(cntr < 6)
  {
        # is the value of cntr within 1 of currentVal continue function if not exit as no other
    if(vect[cntr] < currentVal+2)
    {
      # is the value the same as current val
      # if yes then set onMultiple to TRUE and multipleLength to 1
      if(vect[cntr] == currentVal)
      {
        # check to see if this is 3rd or 4th card in a row of same val
        if(onMultiple)
        {
          multipleLength<- multipleLength + 1
          currentVal<- vect[cntr]
          cntr<- cntr + 1
          cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
        }
        # otherwise it is a new multiple and the 2 cards need adding to multiple length
        else
        {
          multipleLength<- multipleLength + 2
          onMultiple<- TRUE
          currentVal<- vect[cntr]
          cntr<- cntr + 1
          cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
        }
      }
      # the card is on a potential run
      else
      {
        runLength<- runLength +1
        onMultiple<- FALSE
        currentVal<- vect[cntr]
        cntr<- cntr + 1
        cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
      }
    }
    # else not part of a run
    # check that a run of 3 or greater has not ended as their will be no more runs after this
    #   and the points will need to be calculated
    else if(runLength > 2)
    {
      #if multipleLenght = 0 change it to 1 so it can be multiplied with runLength
      if(multipleLength == 0)
      {
        multipleLength<- 1
      }
      
      return(runLength * multipleLength)
    }
    # reset counters and carry on
    else
    {
      onMultiple<- FALSE
      runLength<- 1
      multipleLength<- 0
      currentVal<- vect[cntr]
      cntr<- cntr + 1
      cycleRuns(vect, cntr, currentVal, runLength, multipleLength, onMultiple)
    }
  }
  # check to see if the hand ended on a run
  else if(runLength > 2)
  {
    #if multipleLenght = 0 change it to 1 so it can be multiplied with runLength
      if(multipleLength == 0)
      {
        multipleLength<- 1
      }
      
      return(runLength * multipleLength)
  }
  else
  {
    return(0)
  }
}


```


# calculateScore

## expected input and output

calculateScore(data.frame( value = c(1,2,3,4,5),
                    suit = c("d","d","d","d","d"),
                    cut = c(FALSE,FALSE,FALSE,FALSE,TRUE),
                    stringsAsFactors = FALSE)) # == 12
calculateScore(data.frame( value = c(5,10,10,10,10),
                    suit = c("d","d","c","s","h"),
                    cut = c(FALSE,FALSE,FALSE,FALSE,TRUE),
                    stringsAsFactors = FALSE)) # == 20
calculateScore(data.frame( value = c(5,5,5,5,10),
                    suit = c("d","c","h","s","d"),
                    cut = c(FALSE,FALSE,FALSE,FALSE,TRUE),
                    stringsAsFactors = FALSE)) # == 28
calculateScore(data.frame( value = c(5,5,5,5,11),
                    suit = c("d","c","h","s","d"),
                    cut = c(TRUE,FALSE,FALSE,FALSE,FALSE),
                    stringsAsFactors = FALSE)) # == 29
                    

```{r}

calculateScore<- function(hand)
{
  points<- count15s(hand)
  points<- points + countRuns(hand)
  points<- points + flushScorer(hand)
  points<- points + pair(hand)
  points<- points + nobs(hand)
  return(points)
}

```



# calculateOneHand

```{r}
#takes 4 cards from the full hand and work out the scores for each cut card
calculateOneHand<- function(chosenHand)
{
  
  # loop for 46 times getting the remaining cards in the pack
  for(i in 1:46)
  {
    # append cut card to hand
    hand<- rbind(chosenHand, pack[i,])
    
    # set cut card 'cut' to TRUE
    hand[5,3]<- TRUE
    
    # order the hand by value and suit
    hand <- hand[order(hand[,1], hand[,2]),]
    
    #pass hand to calculateScore
    points<- calculateScore(hand)
    
    print(points)
  }
  
}

```












